---
title: "TIME SERIES ANALYSIS"
author: "Victor Anton, Imholz Chris"
date: "19.12.2024"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

Structure
## 1. Identification phase: Identification of the appropriate model (descriptive analysis!)0

### Descriptive analysis
Objective: Understand how the time series in question are characterized. Results are used as a basis for adequately modelling the subject in question and hence as a basis for inference analysis and forecasting.

− Visualization of the data using basic graphical methods
− Data preparation and transformation
− Decomposition of time series
− Description of the available data using statistical analysis (central moments of the distribution etc.)
− Multivariate analysis: How do different variables relate to each other?
− Analysis of time series characteristics using statistical test methods

## 2. Estimation phase: Model specification and estimation of model parameters

## 3. Diagnostic phase: Evaluation of the quality and validity of the model

## 4. Operational phase: Use of the estimated model for inference analysis or forecasting


```{r libraries, warning=FALSE, message=FALSE}

# Load libraries

```

```{r titles_of_code, echo = TRUE, fig.width=10}

CODE
```

```{r titles_of_code, echo = TRUE, fig.width=10}

CODE

```

## GenAI Usage