---
title: "The Barney Stinson's 3-Day Rule Stradegy"
author: "Victor Anton, Imholz Chris"
date: "19.12.2024"
output:
  html_document: default
---

# **Enhanced Trading Strategy with the 3-Day Rule**

## **Objective**

The purpose of this report is to design and evaluate a robust **trading strategy** for stocks listed in the **Swiss Market Index (SMI)** using the **3-Day Rule**. The strategy consists of:

1. **Identifying Significant Price Drops**: Recognize stock price drops of ≥ 5%.
2. **3-Day Rule Execution**: Buy the stock 3 days after the significant drop.
3. **Performance Evaluation**: Assess the strategy based on:
   - Returns
   - Sharpe Ratio
   - Cumulative Returns
4. **Exploration of Alternative Strategies**: Implement and compare methods like Mean Reversion and Trend-Following.

1. Identifies significant price drops (≥ 5%).
2. Implements the 3-day rule for buying stocks after such events.
3. Evaluates the performance of the strategy using returns, Sharpe Ratio, and cumulative returns.
4. Enhances analysis with time series decomposition and ARIMA modeling.

---

## **Structure**

### **1. Identification Phase**
- **Visualizing Data**: Using plots to understand stock trends.
- **Preparing Data**: Cleaning and calculating key metrics like percentage price changes.
- **Decomposition**: Splitting stock prices into trend, seasonality, and residuals.
- **Statistical Analysis**: Examining central moments and relationships among variables.

### **2. Estimation Phase**
- **ARIMA Forecasting**: Specifying and estimating time series models.

### **3. Diagnostic Phase**
- **Performance Checks**: Evaluating the model’s accuracy and results using:
  - Returns, Success Rates, and Sharpe Ratios.
  - Backtesting on alternative strategies.

### **4. Operational Phase**
- **Execution and Reporting**: Forecasting future price trends and implementing the trading strategy.

---


## **1. Data Preparation**
### **Load Required Libraries**
The following libraries are necessary for data manipulation, visualization, and modeling:
```{r libraries, message=FALSE, warning=FALSE}
# Install required libraries if not installed
if (!require("quantmod")) install.packages("quantmod")
if (!require("dplyr")) install.packages("dplyr")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("forecast")) install.packages("forecast")
if (!require("TTR")) install.packages("TTR")

library(quantmod)
library(dplyr)
library(ggplot2)
library(forecast)
library(TTR)
```

### **Column Descriptions**
The following column names appear in the downloaded data. `[SYMBOL]` represents the stock ticker:

| **Column**         | **Description**                                                                                |
|--------------------|-----------------------------------------------------------------------------------------------|
| `[SYMBOL].Open`    | Opening price of the stock on a given trading day.                                             |
| `[SYMBOL].High`    | Highest intraday price of the stock.                                                           |
| `[SYMBOL].Low`     | Lowest intraday price of the stock.                                                            |
| `[SYMBOL].Close`   | Closing price of the stock at the end of the trading day.                                      |
| `[SYMBOL].Volume`  | Number of shares traded during the trading day.                                                |
| `[SYMBOL].Adjusted`| Adjusted closing price accounting for dividends, splits, and other adjustments.                |

### **Download Stock Data**

```{r smi_symbols, echo = TRUE, fig.width=10}
smi_symbols <- c(
  "ABBN.SW", "ADEN.SW", "ALC.SW", "CFR.SW", "GEBN.SW", "GIVN.SW",
  "HOLN.SW", "BAER.SW", "KNIN.SW", "LOGN.SW", "LONN.SW", "NESN.SW",
  "NOVN.SW", "PGHN.SW", "ROG.SW", "SIKA.SW", "SOON.SW", "SLHN.SW",
  "SREN.SW", "SCMN.SW", "UHR.SW", "UBSG.SW", "ZURN.SW"
)

# Function to download data for each stock
download_data <- function(symbols, start_date = "2022-01-01", end_date = Sys.Date()) {
  stock_data <- list()
  for (symbol in symbols) {
    tryCatch({
      stock_data[[symbol]] <- getSymbols(symbol, src = "yahoo", 
                                         from = start_date, 
                                         to = end_date, 
                                         auto.assign = FALSE)
      cat("Downloaded data for:", symbol, "\n")
    }, error = function(e) {
      cat("Error downloading data for:", symbol, "\n")
    })
  }
  return(stock_data)
}

# Download data
smi_data <- download_data(smi_symbols)

# Save the data to RDS for future use
saveRDS(smi_data, file = "smi_stock_data.rds")
```

### **Inspect the Data**
Before proceeding, we verify the structure of the data to ensure it includes the required columns (e.g., adjusted close prices, volume).
```{r data_check}
# Confirm the structure of the data
head(smi_data[["NESN.SW"]], 10) # Example: Nestlé stock data
```

## **2. Descriptive Analysis**

### **Time Series Decomposition**
To understand the behavior of stock prices, we decompose them into **trend**, **seasonality**, and **residual components**:
- **Trend**: Long-term movement of prices.
- **Seasonality**: Repeated patterns over time.
- **Residuals**: Unexplained variability after removing trend and seasonality.

```{r decomposition}
# Function to decompose stock prices into trend, seasonality, and residuals
decompose_stock <- function(stock_data, symbol) {
  adjusted_col <- paste0(symbol, ".Adjusted")
  if (adjusted_col %in% colnames(stock_data)) {
    ts_data <- ts(stock_data[, adjusted_col], frequency = 252) # Approx. 252 trading days per year
    decomposition <- stl(ts_data, s.window = "periodic")
    plot(decomposition, main = paste("Decomposition for", symbol))
    return(decomposition)
  } else {
    cat("Adjusted price column not found for:", symbol, "\n")
  }
}

# Decompose Nestlé stock data
decompose_stock(smi_data[["NESN.SW"]], "NESN.SW")
```


---


## **3. ARIMA Modeling**

### **Forecasting with ARIMA**
We use ARIMA to model and forecast stock prices:
- ARIMA automatically identifies and fits the best model for time series data.
- Forecasting provides insights into potential future price movements.

```{r arima_forecast}
# Function to fit ARIMA model and forecast prices
forecast_arima <- function(stock_data, symbol, days_ahead = 30) {
  adjusted_col <- paste0(symbol, ".Adjusted")
  if (adjusted_col %in% colnames(stock_data)) {
    ts_data <- ts(stock_data[, adjusted_col], frequency = 252)
    arima_model <- auto.arima(ts_data)
    forecast_prices <- forecast(arima_model, h = days_ahead)
    plot(forecast_prices, main = paste("ARIMA Forecast for", symbol))
    return(forecast_prices)
  } else {
    cat("Adjusted price column not found for:", symbol, "\n")
  }
}

# Forecast Nestlé stock prices
forecast_arima(smi_data[["NESN.SW"]], "NESN.SW")
```

### **Forecasted Values**
- The blue line shows the predicted stock prices for the next period (e.g., next 30 days if specified in the forecast function).
- The model assumes a continuation of the recent downward trend but stabilizes slightly.

### **Confidence Intervals**

The shaded area (light and dark gray) around the forecast line represents the confidence intervals for the forecast:
- Light Gray Band: 95% confidence interval, which is wider to account for greater uncertainty further into the forecast period.
- Dark Gray Band: 80% confidence interval, indicating a narrower range of predictions.
These intervals widen as the forecast horizon increases, reflecting the increasing uncertainty in long-term predictions.

### **Interpretation**

1. The ARIMA model has detected a downward trend in the historical data and projects a continuation of this trend.
2. The widening of the confidence intervals indicates that while the forecast captures the overall trend, the stock's future movements may vary 3. significantly due to volatility.
The downward slope in the forecast suggests bearish sentiment or continued decline in price if no reversal occurs.


---


## **4. Trading Strategy Implementation**

### **3-Day Rule: Event Identification and Trading**
The 3-Day Rule is implemented as follows:
- Identify significant price drops (≥ 5%).
- Buy the stock 3 days after the event.
- Hold the stock for 30 days.
- Calculate returns.

```{r trading_strategy}
implement_3_day_rule <- function(stock_data, symbol, threshold = -5, hold_days = 30) {
  adjusted_col <- paste0(symbol, ".Adjusted")
  
  # Check if the adjusted price column exists
  if (adjusted_col %in% colnames(stock_data)) {
    stock_data <- as.data.frame(stock_data)
    stock_data$Date <- as.Date(rownames(stock_data))  # Ensure we extract dates from row names
    
    # Calculate daily percentage changes
    stock_data$Percent_Change <- (stock_data[, adjusted_col] / dplyr::lag(stock_data[, adjusted_col]) - 1) * 100
    
    # Identify events: significant drops
    stock_data$Event <- stock_data$Percent_Change <= threshold
    
    # Add Buy_Date: 3 days after the event
    stock_data$Buy_Date <- ifelse(stock_data$Event, stock_data$Date + 3, NA)
    
    # Calculate returns: Hold for 30 days
    stock_data$Return <- NA  # Initialize Return column
    for (i in which(stock_data$Event)) {
      buy_price <- stock_data[i + 3, adjusted_col]  # Price 3 days after the event
      sell_index <- i + 3 + hold_days  # Index for sell price
      if (sell_index <= nrow(stock_data)) {
        sell_price <- stock_data[sell_index, adjusted_col]
        stock_data$Return[i] <- (sell_price - buy_price) / buy_price * 100
      }
    }
    
    return(stock_data)
  } else {
    cat("Adjusted price column not found for:", symbol, "\n")
  }
}


# Apply the strategy to Nestlé stock data
strategy_nestle <- implement_3_day_rule(smi_data[["NESN.SW"]], "NESN.SW")
head(strategy_nestle)
```

## **2. Results Interpretation**

### **Table Description**

| **Column**          | **Description**                                                                 |
|----------------------|-------------------------------------------------------------------------------|
| **NESN.SW.Open**     | Opening price of the stock.                                                  |
| **NESN.SW.High**     | Highest price during the trading day.                                        |
| **NESN.SW.Low**      | Lowest price during the trading day.                                         |
| **NESN.SW.Close**    | Closing price at the end of the trading day.                                 |
| **NESN.SW.Volume**   | Number of shares traded on that day.                                         |
| **NESN.SW.Adjusted** | Adjusted closing price (dividends, splits).                                  |
| **Date**             | Trading date for each record.                                                |
| **Percent_Change**   | Daily percentage change in adjusted closing price.                           |
| **Event**            | `TRUE` if a significant price drop occurred (≥ 5%); otherwise `FALSE`.       |
| **Buy_Date**         | Date to buy the stock, which is 3 days after a significant price drop.       |
| **Return**           | Percentage return after holding the stock for 30 days.                      |

---

### **Initial Results for NESN.SW**

```{r display_results}
# Show the first rows of the strategy table
head(strategy_nestle)
```

### **Observations**

From the table:
1. **Percent_Change**: Daily percentage changes are calculated.  
   - For instance, on `2022-01-05`, the stock price dropped by `-2.70%`.  
   - However, this does **not** trigger the event, as the drop is smaller than the `-5%` threshold.

2. **Event**:  
   - The column remains `FALSE` because no significant price drops (`≤ -5%`) are observed in these rows.

3. **Buy_Date and Return**:  
   - Since no events are detected, there are no buy dates (`NA`) or returns calculated at this stage.

---

## **3. Analysis and Next Steps**

### **Current Findings**
- No significant price drops have been observed in the dataset (`Event == FALSE` for all rows).
- As a result, no buy or sell trades have been triggered, and returns remain `NA`.

---

### **Next Steps**
1. **Monitor for Significant Events**:  
   - Continue processing more rows of stock data to identify days with a `Percent_Change ≤ -5%`.
   
2. **Evaluate Strategy Performance**:  
   - Calculate the mean, median, and success rate of returns once events are detected.
   - Visualize cumulative returns over time.

3. **Sensitivity Analysis** (Optional):  
   - Lower the threshold (e.g., `-4%`) to detect more events and assess the strategy's sensitivity.

---

## **4. Visualizing Stock Price Movements**

### **Plot Adjusted Closing Prices**

```{r plot_prices}
# Plot the adjusted closing prices for NESN.SW
plot(strategy_nestle$Date, strategy_nestle$NESN.SW.Adjusted, type = "l", col = "blue",
     main = "Adjusted Closing Prices for NESN.SW",
     xlab = "Date", ylab = "Adjusted Close Price")
grid()
```

## **5. Performance Evaluation**

### **Evaluate Returns**
We calculate the performance of the strategy using:
- Mean Return
- Median Return
- Success Rate (percentage of trades with positive returns)
- Maximum Drawdown (largest decline in value)

```{r performance_evaluation}
# Calculate summary statistics for returns
evaluate_performance <- function(strategy_data) {
  valid_returns <- na.omit(strategy_data$Return)
  summary_stats <- data.frame(
    Mean_Return = mean(valid_returns),
    Median_Return = median(valid_returns),
    Success_Rate = mean(valid_returns > 0),
    Max_Drawdown = min(valid_returns)
  )
  return(summary_stats)
}

# Evaluate performance for Nestlé
performance_nestle <- evaluate_performance(strategy_nestle)
performance_nestle
```

### **Sharpe Ratio**
The Sharpe Ratio measures risk-adjusted returns:

```{r sharpe_ratio}
# Function to calculate Sharpe Ratio
calculate_sharpe <- function(strategy_data, risk_free_rate = 0.01) {
  valid_returns <- na.omit(strategy_data$Return)
  excess_returns <- valid_returns - risk_free_rate
  sharpe_ratio <- mean(excess_returns) / sd(excess_returns)
  return(sharpe_ratio)
}

# Calculate Sharpe Ratio for Nestlé
sharpe_ratio_nestle <- calculate_sharpe(strategy_nestle)
sharpe_ratio_nestle
```

### **Visualization of Returns**
We visualize cumulative returns over time:

```{r plot_cumulative_retunr}
# Plot cumulative returns
plot_cumulative_returns <- function(strategy_data, symbol) {
  valid_returns <- na.omit(strategy_data$Return)
  cum_returns <- cumsum(valid_returns)
  plot(cum_returns, type = "l", col = "blue",
       main = paste("Cumulative Returns for", symbol),
       xlab = "Trades", ylab = "Cumulative Returns (%)")
}

# Plot cumulative returns for Nestlé
plot_cumulative_returns(strategy_nestle, "NESN.SW")
```

## **Summary**
- **Decomposition**: Time series decomposition reveals trends and seasonality.
- **ARIMA Modeling**: Forecasting provides insights into expected price movements.
- **3-Day Rule**: The strategy calculates buy dates and evaluates holding period returns.
- **Performance Metrics**: Include mean return, Sharpe ratio, and cumulative return visualization.

##################- PIVOTANDO DE ESTRATEGIA #######################################

##Recommendations for Next Steps
1. Relax the Event Threshold
The strict -5% threshold for price drops can be reduced to -3% or -4% to capture more events. This adjustment allows the strategy to identify smaller price drops, which may still offer profitable trading opportunities.

```{r return_visualization}
# Adjust the threshold to -3% or -4% in the function call
strategy_nestle <- implement_3_day_rule(smi_data[["NESN.SW"]], "NESN.SW", threshold = -3)
```

2. Explore Alternative Strategies
Mean Reversion Strategy
The stock's general downward trend might suggest a need for a mean reversion approach:

Identify periods when the stock price deviates significantly from a moving average (e.g., 20-day or 50-day).
Enter a trade when prices are expected to revert back toward the mean.
```{r return_visualization}
# Mean reversion signal: Compare price with 20-day moving average
mean_reversion_signal <- function(stock_data, symbol, ma_period = 20) {
  adjusted_col <- paste0(symbol, ".Adjusted")
  if (adjusted_col %in% colnames(stock_data)) {
    stock_data$MA <- TTR::SMA(stock_data[, adjusted_col], n = ma_period)
    stock_data$Signal <- ifelse(stock_data[, adjusted_col] < stock_data$MA * 0.95, "Buy", "Hold")
    return(stock_data)
  } else {
    cat("Adjusted price column not found for:", symbol, "\n")
  }
}

# Apply mean reversion strategy
mean_reversion_nestle <- mean_reversion_signal(smi_data[["NESN.SW"]], "NESN.SW")
head(mean_reversion_nestle)
```
Trend Following Strategy
Given the observable downward trend, a trend-following strategy might better align with the stock's behavior:

Use indicators such as Simple Moving Average (SMA) or Exponential Moving Average (EMA) to confirm the trend direction.
Trade in the direction of the trend (e.g., short sell during bearish trends).
```{r return_visualization}
# Trend following signal using 50-day EMA
trend_following_signal <- function(stock_data, symbol, ema_period = 50) {
  adjusted_col <- paste0(symbol, ".Adjusted")
  if (adjusted_col %in% colnames(stock_data)) {
    stock_data$EMA <- TTR::EMA(stock_data[, adjusted_col], n = ema_period)
    stock_data$Signal <- ifelse(stock_data[, adjusted_col] < stock_data$EMA, "Short", "Hold")
    return(stock_data)
  } else {
    cat("Adjusted price column not found for:", symbol, "\n")
  }
}

# Apply trend-following strategy
trend_following_nestle <- trend_following_signal(smi_data[["NESN.SW"]], "NESN.SW")
head(trend_following_nestle)
```
3. Perform Backtesting
Implement a backtesting framework to evaluate the performance of the alternative strategies:

Compare cumulative returns and Sharpe Ratios for each strategy.
Assess risk metrics, such as drawdown and volatility.

4. Sensitivity Analysis
Test the strategies with different holding periods (e.g., 20 days instead of 30 days).
Evaluate performance under varying thresholds (e.g., -3%, -4%, -5%).
Summary
Relax the event threshold to -3% or -4% for the 3-Day Rule strategy.
Explore alternative approaches, such as Mean Reversion or Trend Following strategies.
Backtest these strategies to compare performance and refine parameters.
Let me know if you'd like me to integrate these steps directly into your code or continue refining the analysis! 🚀


##################### ESTRATEGIA 5% CON OTROS STOCK ##########################3

Identify Significant 5% Drops
We implement the strict 5% threshold to detect price drops.

```{r return_visualization}
# Function to calculate percent change and identify drops
calc_percent_diff <- function(stock_data, symbol, threshold = -5) {
  adjusted_col <- paste0(symbol, ".Adjusted")
  if (adjusted_col %in% colnames(stock_data)) {
    stock_data <- as.data.frame(stock_data)
    stock_data$Percent_Change <- (stock_data[, adjusted_col] / lag(stock_data[, adjusted_col]) - 1) * 100
    stock_data$Event <- stock_data$Percent_Change <= threshold
    return(stock_data)
  }
}

# Loop through all stocks to identify significant drops
identify_5_percent_drops <- function(data, symbols) {
  results <- list()
  for (symbol in symbols) {
    cat("Processing:", symbol, "\n")
    stock <- data[[symbol]]
    if (!is.null(stock)) {
      stock_with_events <- calc_percent_diff(stock, symbol)
      if (any(stock_with_events$Event, na.rm = TRUE)) {
        results[[symbol]] <- stock_with_events
      }
    }
  }
  return(results)
}

# Apply the function
smi_events <- identify_5_percent_drops(smi_data, smi_symbols)

# Check stocks with 5% downward events
names(smi_events)
```

3. Incorporating Alternative Strategies
Mean Reversion Strategy
We identify stocks that deviate significantly below their 20-day moving average.

```{r return_visualization}
mean_reversion_signal <- function(stock_data, symbol, ma_period = 20) {
  adjusted_col <- paste0(symbol, ".Adjusted")
  if (adjusted_col %in% colnames(stock_data)) {
    stock_data$MA <- SMA(stock_data[, adjusted_col], n = ma_period)
    stock_data$Signal <- ifelse(stock_data[, adjusted_col] < stock_data$MA * 0.95, "Buy", "Hold")
    return(stock_data)
  }
}

# Example for a single stock with mean reversion
if (!is.null(smi_data[["NESN.SW"]])) {
  nesn_mean_reversion <- mean_reversion_signal(smi_data[["NESN.SW"]], "NESN.SW")
  head(nesn_mean_reversion)
}
```

Trend-Following Strategy
We use a 50-day Exponential Moving Average (EMA) as an indicator of trend direction.

```{r return_visualization}
trend_following_signal <- function(stock_data, symbol, ema_period = 50) {
  adjusted_col <- paste0(symbol, ".Adjusted")
  if (adjusted_col %in% colnames(stock_data)) {
    stock_data$EMA <- EMA(stock_data[, adjusted_col], n = ema_period)
    stock_data$Signal <- ifelse(stock_data[, adjusted_col] < stock_data$EMA, "Short", "Hold")
    return(stock_data)
  }
}

# Example for a single stock with trend following
if (!is.null(smi_data[["NESN.SW"]])) {
  nesn_trend_following <- trend_following_signal(smi_data[["NESN.SW"]], "NESN.SW")
  head(nesn_trend_following)
}
```
4. Results and Discussion
Stocks with Significant 5% Drops
The following stocks exhibited at least one significant 5% drop:

```{r return_visualization}
names(smi_events)
```
Visualizing Adjusted Close Prices

```{r return_visualization}
plot_stock <- function(stock_data, symbol) {
  adjusted_col <- paste0(symbol, ".Adjusted")
  if (adjusted_col %in% colnames(stock_data)) {
    plot(stock_data[, adjusted_col], type = "l", col = "blue",
         main = paste("Adjusted Closing Prices for", symbol),
         xlab = "Date", ylab = "Adjusted Price")
  }
}

# Example visualization for a stock with 5% drop
if ("UBSG.SW" %in% names(smi_events)) {
  plot_stock(smi_events[["UBSG.SW"]], "UBSG.SW")
}
```
5. Summary
3-Day Rule Findings:

Significant downward events have been identified for specific stocks (e.g., UBSG.SW).
Returns will be further evaluated.
Alternative Strategies:

Mean Reversion and Trend-Following strategies were introduced for comparison.
Next Steps:

Evaluate returns for identified events and compare performance across strategies.
Perform backtesting to assess historical profitability.
This report sets the stage for refining trading strategies and incorporating robust backtesting frameworks.

