---
title: "The Barney Stinson's 3-Day Rule Strategy"
author: "Victor Anton, Imholz Chris"
date: "19.12.2024"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
    toc: yes
    toc_depth: 1
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: 3
knit:
  citations: yes
---

```{=html}
<style>
body {font-size: 14px;}
p {font-size: 14px;}
h1{font-size: 17pt;}
h2{font-size: 15pt;}
h3,h4,h5,h6{font-size: 14pt;}
div.r-plot {width: 60%; margin: auto;}
img {
  max-width: 60%; 
  height: auto;
  display: block;
  margin: auto;
}
table {
  font-size: 16px;
  line-height: 0.8;
}
</style>

```


```{r setup, include = FALSE, warning = FALSE}
library(knitr)
opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  warning = FALSE,
  message = FALSE)
```

<br>

## **Objective**

The purpose of this report is to design and evaluate a robust **trading strategy** for stocks listed in the **Swiss Market Index (SMI)** using the **3-Day Rule**. The strategy consists of:

1. **Identifying Significant Price Drops**: Recognize stock price drops of ≥ 5%.
2. **3-Day Rule Execution**: Buy the stock 3 days after the significant drop.
3. **Performance Evaluation**: Assess the strategy based on:
   - Returns
   - Sharpe Ratio
   - Cumulative Returns
4. **Exploration of Alternative Strategies**: Implement and compare methods like Mean Reversion and Trend-Following.

1. Identifies significant price drops (≥ 5%).
2. Implements the 3-day rule for buying stocks after such events.
3. Evaluates the performance of the strategy using returns, Sharpe Ratio, and cumulative returns.
4. Enhances analysis with time series decomposition and ARIMA modeling.

---

## **Structure**

### **1. Identification Phase**
- **Visualizing Data**: Using plots to understand stock trends.
- **Preparing Data**: Cleaning and calculating key metrics like percentage price changes.
- **Decomposition**: Splitting stock prices into trend, seasonality, and residuals.
- **Statistical Analysis**: Examining central moments and relationships among variables.

### **2. Estimation Phase**
- **ARIMA Forecasting**: Specifying and estimating time series models.

### **3. Diagnostic Phase**
- **Performance Checks**: Evaluating the model’s accuracy and results using:
  - Returns, Success Rates, and Sharpe Ratios.
  - Backtesting on alternative strategies.

### **4. Operational Phase**
- **Execution and Reporting**: Forecasting future price trends and implementing the trading strategy.

---

## **1. Data Preparation**
### **Load Required Libraries**
## Libraries

In this report the following libraries are used.

<details>

<summary>*Click to see all libraries*</summary>

```{r libraries, class.source = "fold-show"}
if (!require("quantmod")) install.packages("quantmod")
if (!require("dplyr")) install.packages("dplyr")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("forecast")) install.packages("forecast")
if (!require("TTR")) install.packages("TTR")

library(quantmod)
library(dplyr)
library(forecast)
library(TTR)
```

</details>

<br>

### **Column Descriptions**

The following column names appear in the downloaded data. `[SYMBOL]` represents the stock ticker:

| **Column**         | **Description**                                                                                |
|--------------------|------------------------------------------------------------------------------------------------|
| `[SYMBOL].Open`    | Opening price of the stock on a given trading day.                                             |
| `[SYMBOL].High`    | Highest intraday price of the stock.                                                           |
| `[SYMBOL].Low`     | Lowest intraday price of the stock.                                                            |
| `[SYMBOL].Close`   | Closing price of the stock at the end of the trading day.                                      |
| `[SYMBOL].Volume`  | Number of shares traded during the trading day.                                                |
| `[SYMBOL].Adjusted`| Adjusted closing price accounting for dividends, splits, and other adjustments.                |


This section outlines the process of retrieving historical stock price data for companies listed in the Swiss Market Index (SMI) and the S&P 500 index. It specifies the relevant stock symbols and employs a custom function to download the data from Yahoo Finance. The function also includes error handling to manage any issues during the download process. Once the data is downloaded, it is saved as an RDS file to ensure efficient future access and prevent redundant downloads.


```{r shares_download, echo=FALSE, warning = FALSE, fig.width=10}

# Define a list of stock symbols for the Swiss Market Index (SMI). 
# These symbols represent the 20 largest publicly traded companies in Switzerland.
smi_symbols <- c(
  "ABBN.SW", "ADEN.SW", "ALC.SW", "CFR.SW", "GEBN.SW", "GIVN.SW",
  "HOLN.SW", "BAER.SW", "KNIN.SW", "LOGN.SW", "LONN.SW", "NESN.SW",
  "NOVN.SW", "PGHN.SW", "ROG.SW", "SIKA.SW", "SOON.SW", "SLHN.SW",
  "SREN.SW", "SCMN.SW", "UHR.SW", "UBSG.SW", "ZURN.SW"
)


# Retrieve the list of symbols for companies in the S&P 500 index.
# The tq_index function fetches up-to-date information on all S&P 500 components.
# sp500 <- tq_index("SP500")
# head(sp500)

# Function to download historical stock price data for a given list of symbols.
download_data <- function(symbols, start_date = "2022-01-01",end_date = Sys.Date()) {
  stock_data <- list()
  for (symbol in symbols) {
    tryCatch({
      stock_data[[symbol]] <- getSymbols(symbol, src = "yahoo", 
                                         from = start_date, 
                                         to = end_date, 
                                         auto.assign = FALSE)
      cat("Downloaded data for:", symbol, "\n")
    }, error = function(e) {
      cat("Error downloading data for:", symbol, "\n")
    })
  }
  return(stock_data)
}

# Use the download_data function to retrieve historical stock data for all SMI symbols.
smi_data <- download_data(smi_symbols)
#s_p_data <- download_data(sp500$symbol)

# Save the downloaded SMI stock data as an RDS file.
#saveRDS(smi_data, file = "smi_stock_data.rds")
#saveRDS(s_p_data, file = "s_p_stock_data.rds")

```

### **Visualizing Adjusted Stock Prices Over Time**

This section introduces a function for visualizing the adjusted stock prices of specific companies over time. The function ensures that the required adjusted price data is available and generates a line chart displaying price trends. The chart is designed with clear labels, titles, and formatting to enhance readability. This visualization aids in understanding the historical performance of the stocks and identifying trends or patterns that could inform investment strategies.

In this example, we plot the UBS stock prices.

```{r ubs_stock_plot, echo=FALSE, warning = FALSE, fig.width=10}


# Define a custom function `plot_stock` to visualize the adjusted closing prices of a stock.
plot_stock <- function(stock_data, symbol) {
  if (!is.null(stock_data[[symbol]])) {
    stock <- stock_data[[symbol]]
    # Ensure the adjusted price column is available
    column_name <- paste0(symbol, ".Adjusted")
    if (column_name %in% colnames(stock)) {
      plot(stock[, column_name], 
           main = paste("Stock Prices for", symbol), 
           xlab = "Date", 
           ylab = "Adjusted Price", 
           col = "blue", 
           type = "l")
    } else {
      cat("Adjusted price data not available for:", symbol, "\n")
    }
  } else {
    cat("No data available for:", symbol, "\n")
  }
}


# Example usage: Visualize adjusted prices for the UBS stock.
plot_stock(smi_data, "UBSG.SW")

# Example usage: Visualize adjusted prices for Apple stock in the S&P 500.
# plot_stock(s_p_data, "AAPL")
```

### **Calculating Daily Percentage Price Changes for SMI Stocks**

This section calculates the daily percentage price changes for each stock's adjusted closing price in the Swiss Market Index (SMI). A utility function is applied across all stocks to compute these changes and store them in a new column. This transformation prepares the data for identifying significant events, such as large price changes, which are essential for strategy development.

<details>

<summary>*Click to see percentage difference data for "UBSG.SW"*</summary>

```{r percent_diff, class.source = "fold-show"}
# The `calc_percent_diff` function calculates the percentage change in the adjusted closing price for a given stock.
calc_percent_diff <- function(stock_data, symbol) {
  if (!is.null(stock_data[[symbol]])) {
    
    close_col <- paste0(symbol, ".Adjusted")
    if (close_col %in% colnames(stock_data[[symbol]])) {
      
      close_prices <- stock_data[[symbol]][, close_col]
      percent_diff <- (close_prices / lag(close_prices) - 1) * 100
      
      percent_diff_col <- paste0(symbol, ".Percent_Diff")
      
      stock_data[[symbol]] <- as.data.frame(stock_data[[symbol]])
      
      stock_data[[symbol]][, percent_diff_col] <- percent_diff
      
      
      
      return(stock_data[[symbol]])
    } else {
      cat("Close column not found for:", symbol, "\n")
      return(NULL)
    }
  } else {
    cat("No data available for:", symbol, "\n")
    return(NULL)
  }
}

# The `calc_all_percent_diff` function applies the `calc_percent_diff` function to a list of stock symbols.
calc_all_percent_diff <- function(stock_data, symbols) {
  modified_stock_data <- stock_data  # Copy the original list structure
  for (symbol in symbols) {
    modified_stock_data[[symbol]] <- calc_percent_diff(stock_data, symbol)
  }
  return(modified_stock_data)
}

# Calculate percentage differences for all stocks in the SMI.
smi_percent_diff <- calc_all_percent_diff(smi_data, smi_symbols)

# Example: View the first few rows of percentage difference data for "UBSG.SW".
head(smi_percent_diff[["UBSG.SW"]])

```

</details>

<br>



### Implementation of the Strategy

The strategy is implemented in several steps to identify and act on
significant stock price movements systematically:

1.  **Identification of Events**: Significant events are identified as
    price changes of 5% or more in the adjusted closing price.
2.  **Stock Purchase**: Stocks are purchased three trading days after a
    significant event to mitigate overreaction.
3.  **Holding Period**: The stocks are held for 30 trading days,
    allowing for potential recovery.
4.  **Return Calculation**: Returns are calculated based on the buy
    price and the sell price after 30 days.

Each step is supported by specific code implementations, which are
detailed below.

1.  **Identifying Events with ≥ 5% Price Changes**\
    Significant price movements, defined as changes of 5% or more in the
    adjusted closing price, are critical for triggering trading actions.
    To identify such events, we calculate the daily percentage change
    for each stock using the adjusted closing prices. The percentage
    change is computed using the formula:

$$
  \text{Percentage Change} = \left(\frac{\text{Current Price}}{\text{Previous Price}} - 1\right) \times 100
$$

    This calculation is applied across all stocks in the dataset, and
    the results are stored in a new column, `[SYMBOL].Percent_Diff`. A
    threshold filter is then applied to flag events where the percentage
    change is lower or equal to -5%.

<details>

<summary>*Click to see all identified events for UBS*</summary>

```{r event_identification, class.source = "fold-show"}
# 1. Identifying Events with ≥ 5% Price Change.
identify_events <- function(stock_data, symbol, threshold) {
  if (!is.null(stock_data[[symbol]])) {
    percent_diff_col <- paste0(symbol, ".Percent_Diff")
    if (percent_diff_col %in% colnames(stock_data[[symbol]])) {
      stock_data[[symbol]] <- as.data.frame(stock_data[[symbol]])
      stock_data[[symbol]]$Event <- stock_data[[symbol]][, percent_diff_col] < threshold
      return(stock_data[[symbol]])
    } else {
      cat("Percent diff column not found for:", symbol, "\n")
      return(NULL)
    }
  } else {
    cat("No data available for:", symbol, "\n")
    return(NULL)
  }
}

# Initialize an empty list to store results
smi_with_events <- list()

# For loop to identify events
for (i in seq_along(smi_percent_diff)) {
  symbol <- smi_symbols[i]  # Current symbol
  smi_with_events[[symbol]] <- identify_events(stock_data = smi_percent_diff, 
                                               symbol = symbol, 
                                               threshold = -5.0)
}

head(smi_with_events[["UBSG.SW"]], 100)


```

</details>

<br>


#### 2. Buying the Stock 3 Days After the Event

To mitigate market overreaction and ensure a more stable entry point, the strategy specifies buying the stock three trading days after the identified event. This delay allows the market to stabilize and provides an opportunity to enter at a more predictable price level. The corresponding buy date is calculated for each flagged event.

<details>

<summary>*Click to see all libraries*</summary>

```{r buy_3_days_later, class.source = "fold-show"}
#library(dplyr)  # Lade dplyr für if_else
# Funktion zum Hinzufügen des Kaufdatums (3. Handelsdatum nach Event)
add_buy_dates <- function(stock_data, symbol) {
  if (!is.null(stock_data[[symbol]])) {
    stock_data[[symbol]] <- as.data.frame(stock_data[[symbol]])
    
    # TODO : braucht es eventuell nicht mehr
    # Konvertiere den Index in ein Date-Objekt und speichere es als Event_Date
    stock_data[[symbol]]$Event_Date <- as.Date(row.names(stock_data[[symbol]]))
    
    # Initialisiere Buy_Date mit NA
    stock_data[[symbol]]$Buy_Date <- as.Date(NA)
    
    # Berechne das 3. Handelsdatum nach dem Event
    for (i in which(stock_data[[symbol]]$Event)) {
      current_index <- i
      future_index <- current_index + 3  # Das 3. Handelsdatum (Index-basiert)
      
      # Prüfe, ob der zukünftige Index innerhalb der DataFrame-Grenzen liegt
      if (future_index <= nrow(stock_data[[symbol]])) {
        stock_data[[symbol]]$Buy_Date[current_index] <- stock_data[[symbol]]$Event_Date[future_index]
      }
    }
    
    return(stock_data[[symbol]])
  } else {
    cat("No data available for:", symbol, "\n")
    return(NULL)
  }
}

# For-Loop, um Kaufdaten und Ereignisdatum hinzuzufügen
smi_with_buy_dates <- list()

for (i in seq_along(smi_with_events)) {
  symbol <- smi_symbols[i]  # Das aktuelle Symbol
  smi_with_buy_dates[[symbol]] <- add_buy_dates(stock_data = smi_with_events, symbol = symbol)
}

# Beispiel: UBSG.SW anzeigen
head(smi_with_buy_dates[["UBSG.SW"]][, c("Event_Date", "Buy_Date", "Event")], 1000)
```

</details>

<br>


#### 3. Holding the Stock for 30 Days

After the stock is purchased, it is held for a fixed period of 30 trading days. This duration is chosen based on the hypothesis that prices will recover following the initial overreaction, allowing for potential capital appreciation during this window.

#### Calculating Returns

To evaluate the strategy's performance, the return on each trade is calculated as:

$$
  \text{Return} = \frac{\text{Sell Price} - \text{Buy Price}}{\text{Buy Price}}
$$

Where:

- The **Buy Price** is the adjusted closing price on the calculated buy date.
- The **Sell Price** is the adjusted closing price 30 days after the buy date.

These returns are stored for analysis and aggregated to assess the overall profitability of the strategy.

<details>

<summary>*Click to see all libraries*</summary>

```{r holding_period, class.source = "fold-show"}
# Funktion zur Berechnung der Renditen
calculate_returns <- function(stock_data, symbol) {
  if (!is.null(stock_data[[symbol]])) {
    stock_data[[symbol]] <- as.data.frame(stock_data[[symbol]])
    adjusted_col <- paste0(symbol, ".Adjusted")  # Angepasster Schlusskurs

    if (adjusted_col %in% colnames(stock_data[[symbol]])) {
      # Rendite berechnen
      stock_data[[symbol]]$Return <- ifelse(
        !is.na(stock_data[[symbol]]$Buy_Date),  # Wenn ein Kaufdatum existiert
        {
          # Kaufpreis (Schlusskurs am Buy_Date)
          buy_price <- lag(stock_data[[symbol]][, adjusted_col], 3)
          # Verkaufspreis (Schlusskurs nach 30 Tagen)
          sell_price <- lead(stock_data[[symbol]][, adjusted_col], 30)
          # Rendite berechnen
          (sell_price - buy_price) / buy_price
        },
        NA  # Keine Rendite, wenn kein Kaufdatum vorhanden ist
      )
      return(stock_data[[symbol]])
    } else {
      cat("Adjusted price column not found for:", symbol, "\n")
      return(NULL)
    }
  } else {
    cat("No data available for:", symbol, "\n")
    return(NULL)
  }
}

# For-Loop, um Renditen für jedes Symbol zu berechnen
smi_with_returns <- list()

for (i in seq_along(smi_with_buy_dates)) {
  symbol <- smi_symbols[i]  # Das aktuelle Symbol
  smi_with_returns[[symbol]] <- calculate_returns(stock_data = smi_with_buy_dates, symbol = symbol)
}

head(smi_with_returns[["UBSG.SW"]],1000)
```

</details>

<br>


#### 4. Summary of the Results

The performance of the strategy is quantified using several key metrics:

- **Average Return**: The average return across all transactions.
- **Median Return**: The median return, which helps avoid distortion due to outliers.
- **Positive vs. Negative Returns**: The proportion of positive returns relative to negative returns.
- **Maximum and Minimum Return**: Identification of the maximum and minimum returns across the trades.
- **Total Return**: The total return when applying the strategy to the entire portfolio.

These metrics are calculated using the following dataset, which includes event dates, buy dates, and returns:

| Event_Date | Buy_Date   | Return     |
|------------|------------|------------|
| 2022-01-03 | 2022-01-03 | <NA>       |
| 2022-01-04 | 2022-01-04 | <NA>       |
| 2022-01-05 | 2022-01-05 | <NA>       |
| ...        | ...        | ...        |
| 2022-02-24 | 2022-02-24 | -0.0597    |
| 2022-03-01 | 2022-03-01 |  0.0585    |
| 2022-03-04 | 2022-03-04 |  0.0813    |
| ...        | ...        | ...        |

The returns are calculated by comparing the buy and sell prices, with the strategy's overall performance aggregated to assess profitability.

```{r calculate_returns, echo = FALSE, fig.width=10}
# Function to summarize the results and extract returns
summarize_results <- function(stock_data) {
  # Extract all returns from the list
  returns <- do.call(rbind, lapply(stock_data, function(df) {
    if (!is.null(df$Return)) {
      return(df$Return)
    } else {
      return(NULL)
    }
  }))
  
  # Remove NA values
  returns <- returns[!is.na(returns)]
  
  # Calculate metrics
  summary <- list(
    average_return = mean(returns),  # Average return
    median_return = median(returns),  # Median return
    positive_ratio = sum(returns > 0) / length(returns),  # Proportion of positive returns
    max_return = max(returns),  # Maximum return
    min_return = min(returns),  # Minimum return
    total_return = sum(returns)  # Total return
  )
  
  # Store returns globally if needed
  assign("returns", returns, envir = .GlobalEnv)
  
  return(summary)
}

# Summarize results for SMI and make returns globally available
results_summary <- summarize_results(smi_with_returns)

# Print the results
print(results_summary)
```

### 2. **Visualization of Strategy Performance**
This section focuses on visualizing the returns generated by the implemented strategy to gain insights into its performance. Two key visualizations are included:


**a) Distribution of Returns**
Objective:
The histogram shows the frequency of positive and negative returns, providing a clear overview of the return distribution. This helps identify whether the strategy generates more positive or negative returns and highlights the variability of returns.


```{r Distribution of Returns, echo = TRUE, fig.width=10}
ggplot(data = data.frame(Return = returns), aes(x = Return)) +
  geom_histogram(binwidth = 0.01, fill = "blue", color = "white") +
  labs(title = "Distribution of Returns", x = "Return", y = "Frequency") + 
  theme_minimal()

```

**b) Returns Over Time**
Objective:
The scatterplot shows how returns evolve over time, highlighting any trends, clusters, or patterns in the strategy’s performance.


```{r visualize_returns, echo = TRUE, fig.width=10}
# Combine all data frames into a single DataFrame and retain only valid rows
all_returns <- do.call(rbind, lapply(smi_with_returns, function(df) {
  if (!is.null(df$Buy_Date) & !is.null(df$Return)) {
    # Select rows where Buy_Date and Return are not NA
    valid_rows <- df[!is.na(df$Buy_Date) & !is.na(df$Return), c("Buy_Date", "Return")]
    return(valid_rows)
  }
  return(NULL)
}))

# Ensure Buy_Date is treated as a Date object
all_returns$Buy_Date <- as.Date(all_returns$Buy_Date)

# Scatterplot of returns over time with color-coded points in a blue scale
ggplot(data = all_returns, aes(x = Buy_Date, y = Return, color = Return)) +
  geom_point(alpha = 0.7) +  # Semi-transparent points for better visibility
  scale_color_gradient2(low = "lightblue", mid = "blue", high = "darkblue", midpoint = 0) +  # Blue scale
  labs(title = "Returns Over Time", x = "Buy Date", y = "Return") +
  theme_minimal()            # Apply a clean, minimal theme
```

### **Implementation and Benchmarking of the Buy-and-Hold Strategy**

This section focuses on implementing the **Buy-and-Hold strategy** as a benchmark for evaluating the trading strategy. For each stock in the SMI, the strategy calculates the return by comparing the adjusted closing price at the start date to the adjusted closing price at the end date. The returns are averaged across all stocks to provide an overall benchmark return.

It is important to note that the average return calculated here is **not weighted**, unlike the SMI index, which is capitalization-weighted. This means that all stocks are treated equally in the calculation, regardless of their market value.

The Buy-and-Hold approach assumes no active trading decisions, offering a simple and passive investment strategy. By comparing this benchmark to the cumulative return of the trading strategy, we can assess whether active trading outperforms or underperforms a passive investment approach.


```{r benchmark_buy_and_hold, echo = TRUE, fig.width=10}

# Function to implement the Buy-and-Hold strategy
buy_and_hold_strategy <- function(stock_data, symbol, start_date, end_date) {
  if (!is.null(stock_data[[symbol]])) {
    stock_data[[symbol]] <- as.data.frame(stock_data[[symbol]])
    
    adjusted_col <- paste0(symbol, ".Adjusted")  # Adjusted closing price
    
    if (adjusted_col %in% colnames(stock_data[[symbol]])) {
      # Filter data for the specified date range
      stock_subset <- stock_data[[symbol]][as.Date(row.names(stock_data[[symbol]])) >= as.Date(start_date) &
                                             as.Date(row.names(stock_data[[symbol]])) <= as.Date(end_date), ]
      
      # Ensure there is enough data
      if (nrow(stock_subset) >= 2) {
        # Buy price (adjusted closing price on the start date)
        buy_price <- stock_subset[1, adjusted_col]
        # Sell price (adjusted closing price on the end date)
        sell_price <- stock_subset[nrow(stock_subset), adjusted_col]
        # Calculate return
        return <- (sell_price - buy_price) / buy_price
        return(return)
      } else {
        cat("Not enough data for:", symbol, "\n")
        return(NA)
      }
    } else {
      cat("Adjusted price column not found for:", symbol, "\n")
      return(NA)
    }
  } else {
    cat("No data available for:", symbol, "\n")
    return(NA)
  }
}

# Benchmarking for all SMI stocks
benchmark_smi <- function(stock_data, symbols, start_date, end_date) {
  returns <- sapply(symbols, function(symbol) {
    buy_and_hold_strategy(stock_data, symbol, start_date, end_date)
  })
  # Average return of the benchmark
  avg_return <- mean(returns, na.rm = TRUE)
  return(list(individual_returns = returns, average_return = avg_return))
}

# Define the start and end dates for the Buy-and-Hold strategy
start_date <- "2022-01-01"
end_date <- "2024-12-19"

# Calculate the benchmark return for the SMI
benchmark_results <- benchmark_smi(smi_data, smi_symbols, start_date, end_date)

# Display results
cat("Individual returns for SMI stocks:\n")
print(benchmark_results$individual_returns)  # Individual returns of the stocks
cat("\nAverage benchmark return (Buy-and-Hold):", benchmark_results$average_return, "\n")
cat("\nAverage return of SMI stocks (overall):", mean(benchmark_results$individual_returns), "\n")

```

### **Benchmark Comparison**

This section compares the **cumulative return** of the trading strategy to the **average Buy-and-Hold benchmark**. A bar chart visualizes the performance difference, highlighting the strategy's effectiveness against a passive investment approach.


```{r benchmark, echo = TRUE, fig.width=10}  
# Cumulative return of the trading strategy
cumulative_strategy_return <- prod(1 + returns, na.rm = TRUE) - 1  # Convert to percentage
cat("Cumulative Trading Strategy Return:", cumulative_strategy_return * 100, "%\n")

# Average return of the Buy-and-Hold benchmark (already calculated)
average_benchmark_return <- benchmark_results$average_return
cat("Average Benchmark (Buy-and-Hold) Return:", average_benchmark_return * 100, "%\n")

# Comparison visualization
comparison_data <- data.frame(
  Strategy = c("Trading Strategy", "Buy-and-Hold Benchmark"),
  Return = c(cumulative_strategy_return, average_benchmark_return)
)

# Bar plot to compare the returns with a blue color theme
ggplot(comparison_data, aes(x = Strategy, y = Return, fill = Strategy)) +
  geom_bar(stat = "identity", color = "black", alpha = 0.7) +
  labs(title = "Comparison of Trading Strategy vs Buy-and-Hold Benchmark",
       x = "Strategy", y = "Return") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("blue", "lightblue")) +  # Blue theme colors
  theme_minimal() +
  theme(legend.position = "none")  # Hide the legend
```

## **Performance Analysis: Trading Strategy vs Buy-and-Hold Benchmark**

The results and visualization clearly show that the **trading strategy** significantly underperformed the **Buy-and-Hold benchmark**. The following is a detailed analysis and interpretation:

---

### **1. Trading Strategy Performance**
- **Cumulative Trading Strategy Return**: **-99.99%**
   - This indicates that the trading strategy almost completely wiped out the investment, leading to an extreme loss.
   - A return of **-99.99%** implies that the capital has been almost entirely depleted, leaving very little value.

---

### **2. Buy-and-Hold Benchmark Performance**
- **Average Benchmark Return**: **+9.67%**
   - The **Buy-and-Hold strategy** generated a positive return of **9.67%** over the same period.
   - This suggests that holding SMI stocks throughout the time horizon would have resulted in moderate but consistent gains, far outperforming the active strategy.

---

### **3. Key Insights**
- The **trading strategy** suffered from significant underperformance compared to the Buy-and-Hold benchmark.
- A **-99.99% cumulative return** points to fundamental issues with the strategy, possibly due to:
   - **Ineffective timing** of entries and exits.
   - **Inadequate management** of major price drops or trends.
   - The strategy might have consistently flagged stocks that were on a downward trajectory, resulting in poor trade decisions.
- In contrast, the **Buy-and-Hold benchmark** delivered steady returns, suggesting that a passive investment strategy in SMI stocks could have outperformed the active trading approach over the same period.

---

### **4. Potential Causes of the Strategy's Failure**
The catastrophic underperformance of the strategy may be attributed to several factors:
1. **Event Identification**: The threshold for significant price movements (e.g., -5%) might have led to frequent trades in stocks that were in a prolonged decline rather than capturing potential recoveries.
2. **Holding Period**: A holding period of 30 days may not have been enough to realize gains from rebounds or recoveries, or the market conditions may have continued to decline during that period.
3. **Execution Logic**: There may be flaws in the buy/sell price calculation, especially in how adjusted closing prices or lagged values were incorporated, which could have distorted the performance.
4. **Market Conditions**: A generally bearish or volatile market during the evaluation period may have exacerbated losses, amplifying the underperformance of the strategy.

---

### **5. Recommendations**
To improve the strategy's performance, several adjustments are recommended:
1. **Refine the Strategy**:
   - Reevaluate the event thresholds for identifying significant price changes and adjust entry rules.
   - Experiment with different holding periods based on broader market trends to better capture recoveries or market cycles.
2. **Introduce Stop-Loss Logic**:
   - Implement stop-loss mechanisms to mitigate losses in case of adverse market movements following trade execution.
3. **Validate Data**:
   - Ensure that price and return calculations are accurate, free of missing values, and correctly reflect market conditions.
4. **Backtest the Strategy**:
   - Test the strategy over different time periods and market conditions to assess its robustness and adaptability to various market environments.

---

### **Conclusion**
The severe underperformance of the trading strategy highlights the need for substantial refinements in its design or execution. Meanwhile, the **Buy-and-Hold benchmark** demonstrates that a passive investment strategy in SMI stocks outperformed the active approach. Revising the strategy’s logic and risk management is essential for improving its overall performance.

---

## **GenAI Usage**

In developing and evaluating the trading strategy, **Generative AI (GenAI)** can be instrumental in improving the design and backtesting process. By leveraging AI models, we can automate the identification of optimal entry and exit points, conduct more accurate sentiment analysis, and even simulate different market scenarios to test the strategy's robustness. GenAI tools can also be used to enhance the event identification logic, optimize the holding periods, and suggest dynamic adjustments to risk management parameters. Furthermore, the integration of natural language processing (NLP) models can aid in analyzing market news and investor sentiment, providing additional insights to refine strategy decisions and predict market trends. With these advanced AI techniques, the strategy’s performance could be significantly improved, offering more consistent returns and better risk control.
